---
title: "M06-Reflection Essay-Advanced Data Wrangling"
author: "Nik Sahami"
date: today
format: 
  html: 
    theme: cosmo
    toc: true
    toc-depth: 4
    number-sections: true
    code-link: true
    embed-resources: true
editor: visual
execute: 
  freeze: auto
---

# Essay Prompts

## Revealjs vs PPT

> *In **Step 1**, you learned how to design stunning presentation slides with Quarto. What's your impression of the Revealjs presentation? Describe its capabilities as you learned from the presentations. What are its strengths and weaknesses compared with PPT?*

My impression of Revealjs is that it is a really useful alternative to PowerPoint, especially for presentations that involve code and technical content. I like that it gives you a lot of control over how the presentation looks and works without having to manually edit every slide. You can use themes, transitions, incremental lists, speaker notes, vertical slides, and even interactive tools like zooming and the chalkboard feature. I also think the ability to fold or highlight parts of code is especially useful for presentations where you are explaining your work.

One of the biggest strengths of Revealjs compared with PowerPoint is that the presentation is created through Quarto, so it can be easily updated and reproduced. Changes to fonts and colors can be applied across the whole presentation instead of changing slides one at a time. It also works really well with code and data outputs, which makes it useful for more analytical presentations.

The main weakness is that Revealjs has a much steeper learning curve than PowerPoint. PowerPoint is easier for quickly moving things around and designing slides visually, but Revealjs requires you to work with Markdown coding and sometimes SCSS. I think PPT is still easier for presentations that need a lot of custom visual placement, and I like being able to drag elements around, but Revealjs is stronger when reproducibility and integration with code are more important.

## Presentarion Link

> *In **Step 2**, you were asked to create a Revealjs presentation. **Provide a link to the presentation you published on the web**.*

[Click here to view the presentation](https://01a10805-7236-14ab-c340-17c302cc4d23.share.connect.posit.cloud/)

The presentation is about my CEP project with the Collins College of Hospitality Management. It focuses on the difference in retention between first-time freshmen and transfer students and how factors such as engagement, extracurricular involvement, financial factors, and demographics may help explain that difference. I also included a reproducible table that summarizes the variables I plan to use in the analysis.

I liked making the presentation and it was pretty easy since I am already familiar with the design of a Quarto document. The coolest parts were how it rendered the revealing information and being able to embed data tables and code into the presentation.

## Advanced Wrangling Examples

> *My advanced wrangling lecture in **Step 3** introduced new concepts, including how to join relational databases. Pick three different functions, each from a different category. Describe when to use them and give one coding example along with an output for each function. To clarify, you are expected to do literate programming to weave your narratives with the code.*

::: {.panel-tabset}
### 1. `left_join()`

`left_join()` is useful when you have information stored in two different data frames and want to combine them using a common key. It keeps every row from the first data frame and adds matching information from the second data frame.

For example, I can combine student information with scholarship information using a student ID.

```{r}
library(tidyverse)

students <- tibble(student_id = c(101, 102, 103),
                   name = c("A", "B", "C"))

scholarships <- tibble(student_id = c(101, 103),
                       scholarship = c("Yes", "Yes"))

students |>
  left_join(scholarships, by = "student_id")
```

The output shows that all students remain in the dataset, even if they do not have a matching scholarship record. Itt adds *scholarship* as a new column, and students without a match receive `NA`.

### 2. `across()`

`across()` is useful when the same function needs to be applied to several columns at once. This is especially helpful when summarizing multiple variables without writing the same code repeatedly.

::: {.callout-note}
The data in the following examples are not real and are only being used to show the functions.
:::

```{r}
survey <- tibble(engagement = c(4, 3, 5, 2),
                 satisfaction = c(5, 4, 4, 3),
                 involvement = c(3, 2, 5, 4))

survey |>
  summarize(across(c(engagement, satisfaction, involvement),
                   mean))
```

The output provides the mean for each selected variable in one step.

### 3. `case_when()`

`case_when()` is useful when I want to create or recode a variable based on multiple conditions. It works well when categories depend on ranges or several logical rules.

For example, I can group students into engagement levels based on their engagement score.

```{r}
survey <- survey |>
  mutate(engagement_level = case_when(engagement >= 4 ~ "High",
                                      engagement == 3 ~ "Moderate",
                                      engagement < 3 ~ "Low"))

survey
```

The output adds a new `engagement_level` column and assigns each student to a category based on their score.

:::

# Appendix

[GitHub Repo Link](https://github.com/itsnik1/m06_reflection)

Published gh pages report
